Appendix 1: Richness and abundance analysis

### 2021-11-05 code/ms1_rich_abund.Rmd
updating just in case

---
title: "Abundance/diversity GAM"
output: html_document
---

# 1. Prepare environment
## 1.1 Load packages
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(magick)
library(patchwork)
library(tidymv)
library(vegan)
library(gratia)
library(tidyverse)
library(broom)
library(lubridate)
library(cowplot)
```

## 1.2 Load site data and floral taxonomy data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, 
                   temp.mean, elev.class, elev.class2)) %>% 
  mutate(site = factor(site),
         tree_line = factor(tree_line))

# FL traits
## family-level taxonomy
fl_tax <- read_csv("../data/processed_data/floral_tax.csv") 
```

@@

### Floral survey data
survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
  # Consider only visited plant species (anytime anywhere, not specifically for a given site*date)
  semi_join(net, by = "plant.sp") %>% 
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  group_by(plant.sp) %>%
  mutate(present = case_when(
    max(flower.cover) > 0 ~ TRUE,
    max(flower.cover) == 0 ~ FALSE
  )) %>%
  filter(present == TRUE) # Ignore plants that never had any flower cover
```

## 1.4 Load images
```{r}
BB <- image_read("../ms_1/figures/phylopic_BB.png")
FL <- image_read("../ms_1/figures/phylopic_FL.png")
BBxFL <- image_read("../ms_1/figures/phylopic_BBxFL.png")
```

## 1.5 Define functions
```{r}
gam_tab1 <-function(x1, x2, x3, 
                    name1, name2, name3) {

  tab1.1 <- x1 %>%
    tidy(parametric = TRUE, conf.int = TRUE) %>%
    mutate(class = rep("parametric", n()),
           estimate = round(estimate, 3),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name1, n())
           )

  tab1.2 <- x1 %>%
    tidy(parametric = FALSE) %>%
    mutate(class = rep("smooth", n()),
           edf = round(edf, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name1, n()))

  tab2.1 <- x2 %>%
    tidy(parametric = TRUE, conf.int = TRUE) %>%
    mutate(class = rep("parametric", n()),
           estimate = round(estimate, 3),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name2, n())
           )

  tab2.2 <- x2 %>%
    tidy(parametric = FALSE) %>%
    mutate(class = rep("smooth", n()),
           edf = round(edf, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name2, n())
           )

  tab3.1 <- x3 %>%
    tidy(parametric = TRUE, conf.int = TRUE) %>%
    mutate(class = rep("parametric", n()),
           estimate = round(estimate, 3),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name3, n())
           )

  tab3.2 <- x3 %>%
    tidy(parametric = FALSE) %>%
    mutate(class = rep("smooth", n()),
           edf = round(edf, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name3, n())
           )

  out <- bind_rows(tab1.1, tab1.2, tab2.1, tab2.2, tab3.1, tab3.2) %>%
    select(model, class, term, estimate, edf, statistic, p.value) %>%
    mutate(term = case_when(
      term == "(Intercept)" ~ "Intercept",
      term == "mgmtnone" ~ "mgmt:none",
      term == "mgmtmowing" ~ "mgmt:mowing",
      term == "year2011" ~ "year:2011",
      term == "year2012" ~ "year:2012",
      term == "s(site)" ~ "s(site)",
      term == "s(net.transects)" ~ "s(samples)",
      term == "s(survey.transects)" ~ "s(samples)",
      term == "s(elev.mean)" ~ "s(elevation)",
      term == "s(elev.mean):year2010" ~ "s(elevation:2010)",
      term == "s(elev.mean):year2011" ~ "s(elevation:2011)",
      term == "s(elev.mean):year2012" ~ "s(elevation:2012)"),
      sig = case_when(
        p.value < 0.001 ~ "***",
        p.value >= 0.001 & p.value < 0.01 ~ "**",
        p.value >= 0.01 & p.value < 0.05 ~ "*",
        p.value >= 0.05 ~ ""
      ),
      p.value = ifelse(p.value < 0.001, "<0.001", p.value),
      statistic = ifelse(statistic < 0.001, "<0.001", statistic),
      edf = ifelse(edf < 0.001, "<0.001", edf)) %>%
    flextable() %>%
    merge_v(j = c("model", "class")) %>%
    valign(j = c("model", "class"), valign = "top") %>%
    autofit()
}

@@

  tab1.1 <- x1 %>%
    tidy(parametric = TRUE, conf.int = TRUE) %>%
    mutate(class = rep("parametric", n()),
           estimate = round(estimate, 3),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name1, n())
           )

  tab1.2 <- x1 %>%
    tidy(parametric = FALSE) %>%
    mutate(class = rep("smooth", n()),
           edf = round(edf, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name1, n()))

  tab2.1 <- x2 %>%
    tidy(parametric = TRUE, conf.int = TRUE) %>%
    mutate(class = rep("parametric", n()),
           estimate = round(estimate, 3),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name2, n())
           )

  tab2.2 <- x2 %>%
    tidy(parametric = FALSE) %>%
    mutate(class = rep("smooth", n()),
           edf = round(edf, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name2, n())
           )

  out <- bind_rows(tab1.1, tab1.2, tab2.1, tab2.2) %>%
    select(model, class, term, estimate, edf, statistic, p.value) %>%
    mutate(term = case_when(
      term == "(Intercept)" ~ "Intercept",
      term == "mgmtnone" ~ "mgmt:none",
      term == "mgmtmowing" ~ "mgmt:mowing",
      term == "year2011" ~ "year:2011",
      term == "year2012" ~ "year:2012",
      term == "s(site)" ~ "s(site)",
      term == "s(net.transects)" ~ "s(samples)",
      term == "s(survey.transects)" ~ "s(samples)",
      term == "s(elev.mean)" ~ "s(elevation)",
      term == "s(elev.mean):year2010" ~ "s(elevation:2010)",
      term == "s(elev.mean):year2011" ~ "s(elevation:2011)",
      term == "s(elev.mean):year2012" ~ "s(elevation:2012)"),
      sig = case_when(
        p.value < 0.001 ~ "***",
        p.value >= 0.001 & p.value < 0.01 ~ "**",
        p.value >= 0.01 & p.value < 0.05 ~ "*",
        p.value >= 0.05 ~ ""
      ),
      p.value = ifelse(p.value < 0.001, "<0.001", p.value),
      statistic = ifelse(statistic < 0.001, "<0.001", statistic),
      edf = ifelse(edf < 0.001, "<0.001", edf)) %>%
    flextable() %>%
    merge_v(j = c("model", "class")) %>%
    valign(j = c("model", "class"), valign = "top") %>%
    autofit()
}
```

@@

# 2. Process data
## 2.1 Calculate transect-wise FL and BB abundance and richness
```{r echo=FALSE, include=TRUE, message=FALSE}
# Survey transects per site
survey_transects_siteyear <- survey %>%
  mutate(year = factor(year(date)),
         site = factor(site)) %>%
  group_by(site, year) %>%
  summarize(survey.transects = length(unique(date)))

@@

### Per-transect floral abund by species
fl_abund_sp <- survey %>%
  group_by(site, date, plant.sp) %>% # group by plant.sp*transect
  summarize(fl.abund = sum(flower.cover)) %>% # get plant.sp*transect total flower cover; this step really shouldn't be required, but in a few cases, there is more than one abundance entry for the same species on a given transect; not sure how that happened, but it's in the original data; alternatively, I could drop the dups or average them. 
  group_by(site, date) %>%
  mutate(prop.fl.abund = fl.abund/sum(fl.abund)) %>% # get the proportional flower cover for each species.
  ungroup() %>%
  complete(site, date, plant.sp, fill = list(fl.abund = 0, prop.fl.abund = 0)) %>%
  semi_join(survey, by = c("site", "date")) %>%  # only include site-dates that actually occurred
  unique() %>%
  mutate(year = factor(year(date))) %>%
  left_join(fl_tax)

@@

## 2.2 Calculate aggregate abundance and richness by site-year
```{r}
### Bumble bee richness per site-year
bb_rich_siteyear <- net %>%
  mutate(year = factor(year(date))) %>%
  group_by(site, year) %>%
  summarize(bb.rich = length(unique(bb.sp))) %>%
  left_join(site_data) %>% # add site data
  mutate(mgmt = factor(management),
         site = factor(site),
         transect = factor(transect)) %>%
  select(-management) %>%
  left_join(net_transects_siteyear)

@@

## 3.1 Preliminary visualization
```{r eval=FALSE, include=FALSE}
ggplot(rich_abund_siteyear.long, aes(elev.mean, z, color = group)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_grid(year~metric) +
  theme_classic()

ggplot(rich_abund_site.long, aes(z, fill = group)) +
  geom_histogram() +
  facet_wrap(~metric) +
  theme_classic()

# Exploring some bivariate relations

ggplot(rich_abund_site, aes(fl.rich, bb.rich)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(rich_abund_site, aes(fl.peakabund, fl.rich)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(rich_abund_site, aes(fl.peakabund, bb.peakabund)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(rich_abund_site, aes(fl.peakabund, bb.rich)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(rich_abund_site, aes(survey.transects, net.transects)) +
  geom_point() +
  geom_smooth(method = "lm") 
```

## 3.2 Model richness ~ elevation

@@

### 3.2.1 BB richness
```{r}
# Sanity check visualization
ggplot(bb_rich_siteyear, aes(elev.mean, bb.rich)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))  +
  facet_wrap(~year, ncol = 1) +
  theme_classic()

@@

# Plot elevation effect
gam1.1_plot1 <- ggplot(gam1.1_elev, aes(elev.mean, est)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), # Uncertainty
              fill = "black", alpha = 0.2) +
  # geom_rug(data = semi_join(site_data, net), 
  #          aes(elev.mean, color = tree_line), 
  #          inherit.aes = FALSE) +
  theme_classic() +
  labs(y = "Conditional effect", x = "Elevation [m]", color = "Tree line")

@@

### 3.2.2 FL richness
```{r}
# Sanity check visualization
ggplot(fl_rich_siteyear, aes(elev.mean, fl.rich)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))  +
  facet_wrap(~year, ncol = 1) +
  theme_classic()

@@

# Plot elevation effect
gam2.1_plot1 <- ggplot(gam2.1_elev, aes(elev.mean, est)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), # Uncertainty
              fill = "black", alpha = 0.2) +
  # geom_rug(data = semi_join(site_data, net), 
  #          aes(elev.mean, color = tree_line), 
  #          inherit.aes = FALSE) +
  theme_classic() +
  labs(y = "Conditional effect", x = "Elevation [m]", color = "Tree line")

@@

### 3.2.3 Interaction richness
```{r}
# Sanity check visualization
ggplot(interaction_rich_siteyear, aes(elev.mean, interaction.rich)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))  +
  facet_wrap(~year, ncol = 1) +
  theme_classic()

@@

# Plot elevation effect
gam3.1_plot1 <- ggplot(gam3.1_elev, aes(elev.mean, est)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), # Uncertainty
              fill = "black", alpha = 0.2) +
  # geom_rug(data = semi_join(site_data, net), 
  #          aes(elev.mean, color = tree_line), 
  #          inherit.aes = FALSE) +
  theme_classic() +
  labs(y = "Conditional effect", x = "Elevation [m]", color = "Tree line")

@@

### 3.1.4 Compile plots
```{r}
gam_plot_richness_ <- ( gam1.1_plot / gam2.1_plot / gam3.1_plot ) +
  #plot_annotation(tag_levels = 'A') +
  plot_layout(nrow = 3)

@@

@@

## 3.3 Model peak abundance ~ elevation
We will use the same model structure that we used for the richness models.

### 3.3.1 BB abundance
```{r}
ggplot(bb_abund_siteyear, aes(elev.mean, bb.peakabund)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))  +
  facet_wrap(~year, ncol = 1) +
  theme_classic()

@@

# Plot elevation effect
gam4.1_plot1 <- ggplot(gam4.1_elev, aes(elev.mean, est)) +
  geom_line(linetype = "dashed") +
  geom_ribbon(aes(ymin = lower, ymax = upper), # Uncertainty
              fill = "black", alpha = 0.2) +
  # geom_rug(data = semi_join(site_data, net), 
  #          aes(elev.mean, color = tree_line), 
  #          inherit.aes = FALSE) +
  theme_classic() +
  labs(y = "Conditional effect", x = "Elevation [m]", color = "Tree line")

@@

### 3.3.2 FL abundance
```{r}
ggplot(fl_abund_siteyear, aes(elev.mean, fl.peakabund)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))  +
  facet_wrap(~year, ncol = 1) +
  theme_classic()

@@

# Plot elevation effect
gam5.1_plot1 <- ggplot(gam5.1_elev, aes(elev.mean, est)) +
  geom_line(linetype = "solid") +
  geom_ribbon(aes(ymin = lower, ymax = upper), # Uncertainty
              fill = "black", alpha = 0.2) +
  # geom_rug(data = semi_join(site_data, net), 
  #          aes(elev.mean, color = tree_line), 
  #          inherit.aes = FALSE) +
  theme_classic() +
  labs(y = "Conditional effect", x = "Elevation [m]", color = "Tree line")

@@

### 3.3.3 Compile plots
```{r}
gam_plot_abundance_ <- ( gam4.1_plot / gam5.1_plot) +
  #plot_annotation(tag_levels = 'A') +
  plot_layout(nrow = 2)

@@

@@

### 3.3.4 BB sp-level abundance ~ elevation
For this model, the negative-binomial family works best. Maybe it handles zeros better than the quasipoisson.
```{r}
ggplot(bb_abund_sp_siteyear, aes(elev.mean, bb.peakabund, color = year)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp, scales = "free")

@@

### 3.3.5 Compile tabulated summaries
```{r}
gam_tab.rich <- gam_tab1(x1 = gam1.1, x2 = gam2.1, x3 = gam3.1, 
                   name1 = "Bumble bee richness", 
                   name2 = "Floral richness", 
                   name3 = "Interaction richness")

@@

gam_tab.abund <- gam_tab2(x1 = gam4.1, x2 = gam5.1,
                   name1 = "Bumble bee abundance", 
                   name2 = "Floral abundance")

@@

gam_tab.abund_sp1 <- gam6.1  %>%
  tidy(parametric = TRUE) %>%
  mutate(class = rep("parametric", n()),
         estimate = round(estimate, 3),
         statistic = round(statistic, 3),
         p.value = round(p.value, 3)
         )

@@

### 3.3.6 BB elevation distribution
```{r}
bb_range <- net %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = max(abund)) %>%
  filter(abund > 0) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))

@@

## 3.4 Model sample-wise abundance ~ elevation * time
### 3.4.1 BB abundance
```{r}
gam7.1 <- gam(bb.abund ~ 
                year +
                mgmt +
                s(site, bs = "re") +
                te(yday, elev.mean, bs=c("gp","tp")) +
                te(yday, elev.mean, by = year, bs = c("tp", "tp"), m = 1),
              family = "nb",
              method = "REML",
              select = TRUE,
              data = bb_abund)

@@

### 3.4.2 FL abundance
```{r}
gam8.1 <- gam(fl.abund ~ 
                year +
                mgmt +
                s(site, bs = "re") +
                te(yday, elev.mean, bs=c("gp","tp")) +
                te(yday, elev.mean, by = year, bs = c("tp", "tp"), m = 1),
              family = "quasipoisson",
              method = "REML",
              select = TRUE,
              data = fl_abund)

>>>>

---
title: "Appendix 1: Richness and abundance analysis"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
---

# 1. Prepare environment
## 1.1 Load packages
```{r setup, message=FALSE}
library(mgcv)
library(mgcViz)
library(magick)
library(patchwork)
library(tidymv)
library(vegan)
library(gratia)
library(tidyverse)
library(broom)
library(lubridate)
library(cowplot)
```

## 1.2 Load site data and floral taxonomy data
```{r message=FALSE}
site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, # drop unused columns
                   temp.mean, elev.class, elev.class2)) %>% 
  mutate(site = factor(site),
         tree_line = factor(tree_line))

# family-level taxonomy
fl_tax <- read_csv("../data/processed_data/floral_tax.csv") 
```

@@

### Floral survey data
survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
  # Consider only visited plant species 
  semi_join(net, by = "plant.sp") %>% 
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  group_by(plant.sp) %>%
  mutate(present = case_when(
    max(flower.cover) > 0 ~ TRUE,
    max(flower.cover) == 0 ~ FALSE
  )) %>%
  filter(present == TRUE) # Ignore plants that never had any flower cover
```

## 1.4 Load images
Bombus  
  - Melissa Broussard
  - http://phylopic.org/image/750eeb4f-7ca5-49ff-a814-27a12d0cb13a/ 
  - https://creativecommons.org/licenses/by/3.0/

Asteraceae
  - http://phylopic.org/image/39335c0c-f879-4df6-90d5-e4f65d90d04e/
  - The person who associated a work with this deed has dedicated the work to the public domain by waiving all of his or her rights to the work worldwide under copyright law, including all related and neighboring rights, to the extent allowed by law.

```{r message=FALSE}
BB <- image_read("../ms_1/figures/phylopic_BB.png")
FL <- image_read("../ms_1/figures/phylopic_FL.png")
BBxFL <- image_read("../ms_1/figures/phylopic_BBxFL.png")
```

## 1.5 Define functions
These are helper functions for tabulating the output of GAMs.
```{r message=FALSE}
gam_tab1 <-function(x1, x2, x3, 
                    name1, name2, name3) {

  tab1.1 <- x1 %>%
    tidy(parametric = TRUE, conf.int = TRUE) %>%
    mutate(class = rep("parametric", n()),
           estimate = round(estimate, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           lower95 = round(conf.low, 2),
           upper95 = round(conf.high, 2),
           Model = rep(name1, n())
           )

  tab1.2 <- x1 %>%
    tidy(parametric = FALSE) %>%
    mutate(class = rep("smooth", n()),
           edf = round(edf, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           Model = rep(name1, n()))

  tab2.1 <- x2 %>%
    tidy(parametric = TRUE, conf.int = TRUE) %>%
    mutate(class = rep("parametric", n()),
           estimate = round(estimate, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           lower95 = round(conf.low, 2),
           upper95 = round(conf.high, 2),
           Model = rep(name2, n())
           )

  tab2.2 <- x2 %>%
    tidy(parametric = FALSE) %>%
    mutate(class = rep("smooth", n()),
           edf = round(edf, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           Model = rep(name2, n())
           )

  tab3.1 <- x3 %>%
    tidy(parametric = TRUE, conf.int = TRUE) %>%
    mutate(class = rep("parametric", n()),
           estimate = round(estimate, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           lower95 = round(conf.low, 2),
           upper95 = round(conf.high, 2),
           Model = rep(name3, n())
           )

  tab3.2 <- x3 %>%
    tidy(parametric = FALSE) %>%
    mutate(class = rep("smooth", n()),
           edf = round(edf, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           Model = rep(name3, n())
           )

  out <- bind_rows(tab1.1, tab1.2, tab2.1, tab2.2, tab3.1, tab3.2) %>%
    select(Model, class, term, estimate, lower95, upper95, edf, p.value) %>%
    mutate(term = case_when(
      term == "(Intercept)" ~ "Intercept",
      term == "mgmtnone" ~ "mgmt:none",
      term == "mgmtmowing" ~ "mgmt:mowing",
      term == "year2011" ~ "year:2011",
      term == "year2012" ~ "year:2012",
      term == "s(site)" ~ "s(site)",
      term == "s(net.transects)" ~ "s(samples)",
      term == "s(survey.transects)" ~ "s(samples)",
      term == "s(elev.mean)" ~ "s(elevation)",
      term == "s(elev.mean):year2010" ~ "s(elevation:2010)",
      term == "s(elev.mean):year2011" ~ "s(elevation:2011)",
      term == "s(elev.mean):year2012" ~ "s(elevation:2012)"),
      sig = case_when(
        p.value < 0.001 ~ "***",
        p.value >= 0.001 & p.value < 0.01 ~ "**",
        p.value >= 0.01 & p.value < 0.05 ~ "*",
        p.value >= 0.05 ~ ""
      ),
      p.value = ifelse(p.value < 0.001, "<0.001", p.value),
      #statistic = ifelse(statistic < 0.001, "<0.001", statistic),
      edf = ifelse(edf < 0.001, "<0.001", edf)) %>%
    as_grouped_data(group = "Model") %>%
    as_flextable() %>%
    # merge_v(j = c("Model", "class")) %>%
    # valign(j = c("Model", "class"), valign = "top") %>%
    merge_v(j = "class") %>%
    valign(j = "class", valign = "top") %>%
    autofit() %>%
    fontsize(size = 12,
             j = 1:2) %>%
    fontsize(size = 12,
             part = "header")
}

@@

  tab1.1 <- x1 %>%
    tidy(parametric = TRUE, conf.int = TRUE) %>%
    mutate(class = rep("parametric", n()),
           estimate = round(estimate, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           lower95 = round(conf.low, 2),
           upper95 = round(conf.high, 2),
           Model = rep(name1, n())
           )

  tab1.2 <- x1 %>%
    tidy(parametric = FALSE) %>%
    mutate(class = rep("smooth", n()),
           edf = round(edf, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           Model = rep(name1, n()))

  tab2.1 <- x2 %>%
    tidy(parametric = TRUE, conf.int = TRUE) %>%
    mutate(class = rep("parametric", n()),
           estimate = round(estimate, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           lower95 = round(conf.low, 2),
           upper95 = round(conf.high, 2),
           Model = rep(name2, n())
           )

  tab2.2 <- x2 %>%
    tidy(parametric = FALSE) %>%
    mutate(class = rep("smooth", n()),
           edf = round(edf, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           Model = rep(name2, n())
           )

  out <- bind_rows(tab1.1, tab1.2, tab2.1, tab2.2) %>%
    select(Model, class, term, estimate, lower95, upper95, edf, p.value) %>%
    mutate(term = case_when(
      term == "(Intercept)" ~ "Intercept",
      term == "mgmtnone" ~ "mgmt:none",
      term == "mgmtmowing" ~ "mgmt:mowing",
      term == "year2011" ~ "year:2011",
      term == "year2012" ~ "year:2012",
      term == "s(site)" ~ "s(site)",
      term == "s(net.transects)" ~ "s(samples)",
      term == "s(survey.transects)" ~ "s(samples)",
      term == "s(elev.mean)" ~ "s(elevation)",
      term == "s(elev.mean):year2010" ~ "s(elevation:2010)",
      term == "s(elev.mean):year2011" ~ "s(elevation:2011)",
      term == "s(elev.mean):year2012" ~ "s(elevation:2012)"),
      sig = case_when(
        p.value < 0.001 ~ "***",
        p.value >= 0.001 & p.value < 0.01 ~ "**",
        p.value >= 0.01 & p.value < 0.05 ~ "*",
        p.value >= 0.05 ~ ""
      ),
      p.value = ifelse(p.value < 0.001, "<0.001", p.value),
      #statistic = ifelse(statistic < 0.001, "<0.001", statistic),
      edf = ifelse(edf < 0.001, "<0.001", edf)) %>%
    as_grouped_data(group = "Model") %>%
    as_flextable() %>%
    # merge_v(j = c("Model", "class")) %>%
    # valign(j = c("Model", "class"), valign = "top") %>%
    merge_v(j = "class") %>%
    valign(j = "class", valign = "top") %>%
    autofit() %>%
    fontsize(size = 12,
             j = 1:2) %>%
    fontsize(size = 12,
             part = "header")
}
```

@@

# 2. Process data
## 2.1 Calculate transect-wise FL and BB abundance and richness
```{r message=FALSE}
# Survey transects per site
survey_transects_siteyear <- survey %>%
  mutate(year = factor(year(date)),
         site = factor(site)) %>%
  group_by(site, year) %>%
  summarize(survey.transects = length(unique(date)))

@@

### Per-transect floral abund by species
fl_abund_sp <- survey %>%
  group_by(site, date, plant.sp) %>% # group by plant.sp*transect
  summarize(fl.abund = sum(flower.cover)) %>% # get plant.sp*transect total flower cover; 
                                              # this step really shouldn't be required, 
                                              # but in a few cases, there is more than one 
                                              # abundance entry for the same species on a given transect
  group_by(site, date) %>%
  mutate(prop.fl.abund = fl.abund/sum(fl.abund)) %>% # get the proportional flower cover for each species.
  ungroup() %>%
  complete(site, date, plant.sp, fill = list(fl.abund = 0, prop.fl.abund = 0)) %>%
  semi_join(survey, by = c("site", "date")) %>%  # only include site-dates that actually occurred
  unique() %>%
  mutate(year = factor(year(date))) %>%
  left_join(fl_tax)

@@

## 2.2 Calculate aggregate abundance and richness by site-year
```{r message=FALSE}
### Bumble bee richness per site-year
bb_rich_siteyear <- net %>%
  mutate(year = factor(year(date))) %>%
  group_by(site, year) %>%
  summarize(bb.rich = length(unique(bb.sp))) %>%
  left_join(site_data) %>% # add site data
  mutate(mgmt = factor(management),
         site = factor(site),
         transect = factor(transect)) %>%
  select(-management) %>%
  left_join(net_transects_siteyear)

@@

## 3.1 Model richness ~ elevation

@@

### 3.1.1 BB richness
```{r message=FALSE}
# Sanity check visualization
ggplot(bb_rich_siteyear, aes(elev.mean, bb.rich)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))  +
  facet_wrap(~year, ncol = 1) +
  theme_classic()

@@

# Plot elevation effect
gam1.1_plot1 <- ggplot(gam1.1_elev, aes(elev.mean, est)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), # Uncertainty
              fill = "black", alpha = 0.2) +
  theme_classic() +
  labs(y = "Conditional effect", x = "Elevation [m]", color = "Tree line")

@@

### 3.1.2 FL richness
```{r message=FALSE}
# Sanity check visualization
ggplot(fl_rich_siteyear, aes(elev.mean, fl.rich)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))  +
  facet_wrap(~year, ncol = 1) +
  theme_classic()

@@

# Plot elevation effect
gam2.1_plot1 <- ggplot(gam2.1_elev, aes(elev.mean, est)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), # Uncertainty
              fill = "black", alpha = 0.2) +
  theme_classic() +
  labs(y = "Conditional effect", x = "Elevation [m]", color = "Tree line")

@@

### 3.1.3 Interaction richness
```{r message=FALSE}
# Sanity check visualization
ggplot(interaction_rich_siteyear, aes(elev.mean, interaction.rich)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))  +
  facet_wrap(~year, ncol = 1) +
  theme_classic()

@@

# Plot elevation effect
gam3.1_plot1 <- ggplot(gam3.1_elev, aes(elev.mean, est)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), # Uncertainty
              fill = "black", alpha = 0.2) +
  theme_classic() +
  labs(y = "Conditional effect", x = "Elevation [m]", color = "Tree line")

@@

### 3.1.4 Compile plots
```{r message=FALSE}
gam_plot_richness_ <- ( gam1.1_plot / gam2.1_plot / gam3.1_plot ) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(nrow = 3)

@@

gam_plot_richness_

@@

## 3.2 Model peak abundance ~ elevation
We will use the same model structure that we used for the richness models.

### 3.2.1 BB abundance
```{r message=FALSE}
ggplot(bb_abund_siteyear, aes(elev.mean, bb.peakabund)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))  +
  facet_wrap(~year, ncol = 1) +
  theme_classic()

@@

# Plot elevation effect
gam4.1_plot1 <- ggplot(gam4.1_elev, aes(elev.mean, est)) +
  geom_line(linetype = "dashed") +
  geom_ribbon(aes(ymin = lower, ymax = upper), # Uncertainty
              fill = "black", alpha = 0.2) +
  theme_classic() +
  labs(y = "Conditional effect", x = "Elevation [m]", color = "Tree line")

@@

### 3.2.2 FL abundance
```{r message=FALSE}
ggplot(fl_abund_siteyear, aes(elev.mean, fl.peakabund)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))  +
  facet_wrap(~year, ncol = 1) +
  theme_classic()

@@

# Plot elevation effect
gam5.1_plot1 <- ggplot(gam5.1_elev, aes(elev.mean, est)) +
  geom_line(linetype = "solid") +
  geom_ribbon(aes(ymin = lower, ymax = upper), # Uncertainty
              fill = "black", alpha = 0.2) +
  theme_classic() +
  labs(y = "Conditional effect", x = "Elevation [m]", color = "Tree line")

@@

### 3.2.3 Compile plots
```{r message=FALSE}
gam_plot_abundance_ <- ( gam4.1_plot / gam5.1_plot) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(nrow = 2)

@@

gam_plot_abundance_

@@

### 3.2.4 BB sp-level abundance ~ elevation
For this model, the negative-binomial family works best. Maybe it handles zeros better than the quasipoisson.
```{r message=FALSE}
ggplot(bb_abund_sp_siteyear, aes(elev.mean, bb.peakabund, color = year)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp, scales = "free")

@@

## 3.3 Compile tabulated summaries
```{r message=FALSE}
gam_tab.rich <- gam_tab1(x1 = gam1.1, x2 = gam2.1, x3 = gam3.1, 
                   name1 = "Bumble bees", 
                   name2 = "Flora", 
                   name3 = "Interactions")

@@

gam_tab.abund <- gam_tab2(x1 = gam4.1, x2 = gam5.1,
                   name1 = "Bumble bees", 
                   name2 = "Flora")

@@

gam_tab.abund_sp1 <- gam6.1  %>%
  tidy(parametric = TRUE, conf.int = TRUE) %>%
  mutate(class = rep("parametric", n()),
         estimate = round(estimate, 2),
         statistic = round(statistic, 3),
         p.value = round(p.value, 3),
         lower95 = round(conf.low, 2),
         upper95 = round(conf.high, 2),
         )

@@

## 3.4 BB elevation distribution
```{r message=FALSE}
bb_range <- net %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = max(abund)) %>%
  filter(abund > 0) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))

@@

## 3.5 Model sample-wise abundance ~ elevation * time  
These models are complex and potentially sensitive to sampling patterns; they should be interpreted cautiously.

### 3.5.1 BB abundance
```{r message=FALSE}
gam7.1 <- gam(bb.abund ~ 
                year +
                mgmt +
                s(site, bs = "re") +
                te(yday, elev.mean, bs=c("gp","tp")) +
                te(yday, elev.mean, by = year, bs = c("tp", "tp"), m = 1),
              family = "nb",
              method = "REML",
              select = TRUE,
              data = bb_abund)

@@

### 3.5.2 FL abundance
```{r message=FALSE}
gam8.1 <- gam(fl.abund ~ 
                year +
                mgmt +
                s(site, bs = "re") +
                te(yday, elev.mean, bs=c("gp","tp")) +
                te(yday, elev.mean, by = year, bs = c("tp", "tp"), m = 1),
              family = "quasipoisson",
              method = "REML",
              select = TRUE,
              data = fl_abund)
